---
title: "Introduction to iamse"
description: >
  Learn how to get started with the basics of iamse.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to iamse}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r ReaddataLoadLibraries, message=FALSE, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      cache = FALSE,
                      warning = FALSE,
                      eval = TRUE,
                      error = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      include = TRUE,
                      collapse = TRUE,
                      comment = "#>",
                      fig.show = "hold",
                      fig.width=8, fig.height=7)
```


The goal of iamse package is to offer a flexible framework for management
strategy evaluation. To get started, make sure to install the most recent
package version from GitHub.


## Installation

The current version of *iamse* (v0.0.9) requires R $>= 4.0.0$ and can be
downloaded from github as follows.

::: .iamse-release
```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("tokami/iamse")
```
:::

The development version can be installed with:

::: .iamse-devel
```{r, eval = FALSE}
remotes::install_github("tokami/iamse", ref = "dev")
```
:::



## Basic usage

After successful installation, the package can be loaded into the R environment.

```{r, eval = TRUE, echo = FALSE, fig.alt = ""}
library(iamse)
```


## More information

The [tutorial vignette](tutorial.html) provides a step-by-step guide to using
*iamse* for example data.
