---
title: "MSE with iamse"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MSE with iamse}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  fig.width = 7,
  fig.height = 5
)
set.seed(123)
```

# Introduction

The **iamse** package provides a framework for intra-annual, age-based
operating models for management strategy evaluation (MSE). This vignette
walks through a simple demo:

1. Load example life history information
2. Define simulation and noise settings
3. Estimate stochastic reference points
4. Define and run harvest control rules (HCRs)
5. Inspect results and basic trade-offs

The example uses small settings (few years and replicates) to keep runtime
reasonable.

# Loading the package and example data

```{r}
library(iamse)

## Load example life history information
data("stocklist")

## Check the main life history table
head(stocklist$mlh)
```

We start from the built-in multi-species life-history table `stocklist$mlh`.
For this demo, we add a simple assumption about length variation and then
prepare the data with `check.dat()`.

```{r}
dat0 <- stocklist$mlh

## Assume a coefficient of variation in length
dat0$CVlen <- 0.1

## Check and prepare data for iamse
dat <- check.dat(dat0)
str(dat, 1)
```

# Simulation settings

Next we define the simulation settings with `check.set()`. For the vignette
we use a small number of projection years and replicates.

```{r}
set <- check.set()

## Number of projection years
set$nysim <- 20

## Number of stochastic replicates (kept small for demonstration)
set$nrep <- 10

## Use the median across replicates as reference method
set$refMethod <- "median"
```

## Noise structure

We now specify variation in life history, process, and observation
components. Each entry is given as a three-element vector and interpreted
inside **iamse**.

```{r}
## Between-replicate variation (life-history parameters)
set$noise$rep$H  <- c(0.1, 0, 1)
set$noise$rep$M  <- c(0.1, 0, 1)
set$noise$rep$R0 <- c(0.1, 0, 1)

## Process noise (recruitment and fishing mortality)
set$noise$time$R <- c(0.7, 0.4, 1)   ## species-specific recruitment variability
set$noise$rep$F  <- c(0.15, 0, 1)

## Observation noise (indices and catches)
set$noise$time$I <- c(0.3, 0, 1)
set$noise$time$C <- c(0.1, 0, 1)
```

# Stochastic reference points

Before running the MSE, we estimate stochastic reference points for each
stock. To keep computation time down, we reduce the number of replicates
used for the reference calculation.

```{r}
## Lower number of replicates to speed up reference-point estimation
set$refN <- 100

## Estimate stochastic reference levels
dat <- est.ref.levels.stochastic(
  dat, set,
  fmax = 5,
  ncores = 2,
  plot = TRUE
)

## Add Blim as a fraction of pristine biomass
dat$ref$Blim <- 0.2 * dat$ref$B0

str(dat$ref, 1)
```

# Defining harvest control rules

We now define a few simple HCRs:

- `"noF"`: no fishing
- `"refFmsy"`: fishing at the true \(F_\text{MSY}\)

```{r}
## Baseline HCRs

## No fishing ("noF")
iamse::def.hcr.ref(set. = set)

## Fishing at true Fmsy ("refFmsy")
iamse::def.hcr.ref(consF = "fmsy", set. = set)

## Select HCRs to evaluate
set$hcr <- c(
  "noF",
  "refFmsy"
)

set$hcr
```

# Running the MSE

With data, settings, reference points, and HCRs defined, we can run the
MSE. For the vignette we keep `ncores = 1` to avoid parallelization issues
on some systems.

```{r}
resMSE <- run.mse(
  dat, set,
  ncores  = 1,
  verbose = TRUE
)

## Structure of results: first by HCR, then by replicate
names(resMSE)
names(resMSE[[1]])
```

Each element of `resMSE[[hcr]][[rep]]` contains the full output of the
operating model for that HCR and replicate.

```{r}
## Inspect the structure for the first HCR and first replicate
str(resMSE[[1]][[1]], 1)

## For example, TAC time series for a given HCR and replicate
resMSE[[2]][[4]]$tacs$TAC
```

# Time series plots

The package provides helper plotting functions to visualize the evolution
of fishing mortality, catch, and biomass over time for the different HCRs.

```{r}
par(mfrow = c(3, 1), mar = c(4, 4, 1, 1), oma = c(1, 1, 1, 1))

plotiamse.f(dat, set, resMSE)
plotiamse.cw(dat, set, resMSE)
plotiamse.b(dat, set, resMSE)
```

These plots show how the different management strategies affect the
trajectory of the stock and fishery.

# Trade-offs between performance metrics

Finally, we can summarize MSE results using standard performance metrics
such as:

- Long-term catch
- Risk (e.g. probability of dropping below a biomass limit)
- Interannual catch variability (AAV of catch)

```{r}
mets <- est.metrics(
  resMSE, dat, set,
  mets = c("catch", "risk", "aavc")
)

head(mets)
```

We can then visualize trade-offs between pairs of metrics using
`plotiamse.tradeoff()`.

```{r}
par(mfrow = c(1, 2))

## Catch vs risk
plotiamse.tradeoff(
    mets, c("catch", "risk"),
    plot.legend = FALSE
)

## Catch vs interannual variability (AAVC)
plotiamse.tradeoff(
    mets, c("catch", "aavc")
)
```

These trade-off plots provide a compact way to compare HCRs in terms of
yield, conservation risk, and stability — one of the key strengths of
management strategy evaluation.

# Summary

This vignette showed a minimal workflow with **iamse**:

1. Prepare life-history data and simulation settings
2. Specify noise and estimate stochastic reference points
3. Define and select harvest control rules
4. Run the MSE and inspect outputs
5. Summarize and visualize performance trade-offs

You can build on this template by:

- Changing life-history parameters
- Adding more complex noise structures
- Defining additional HCRs
- Increasing the number of years and replicates for more realistic
  applications.
